Updated Constraints on Self-Interacting Dark Matter from Supernova 1987A

### 2017-01-07 draft.tex
Significant additions and edits to section 3.

\arz{Cameron, can you double check this. I think there were some errors in the momenta and in the relative 
signs of the diagrams. I tried to fixe them.}
The eight diagrams contribute the following eight terms to the pp amplitude, 
%
\begin{widetext}
\begin{eqnarray}
	M_1 &=& \frac{4 M_N}{ m_\pi} \frac{f_{pp}^2 e \varepsilon}{k^2-m_\pi^2}  \frac{1}{m_{A^\prime}^2 - 2 \qa \cdot p_1} 
	\bar{u}(p_4) \gamma_5 u(p_2) \bar{u}(p_3) \gamma_5 (\slashed{p}_1 - \qa +M_N)\slashed \epsilon  u(p_1), \\
	M_2 &=& -\frac{4 M_N}{ m_\pi} \frac{f_{pp}^2 e \varepsilon}{l^2-m_\pi^2}  \frac{1}{m_{A^\prime}^2 - 2 \qa \cdot p_1} 
	\bar{u}(p_3) \gamma_5 u(p_2) \bar{u}(p_4) \gamma_5 (\slashed{p}_1 - \qa +M_N)\slashed{\epsilon} u(p_1),\\
	M_3 &=& \frac{4 M_N}{ m_\pi} \frac{f_{pp}^2 e \varepsilon}{l^2-m_\pi^2}  \frac{1}{m_{A^\prime}^2 - 2 \qa \cdot p_2} 
	\bar{u}(p_3) \gamma_5 u(p_1) \bar{u}(p_4) \gamma_5 (\slashed{p}_2 - \qa +M_N)\slashed \epsilon  u(p_2), \\
	M_4 &=& -\frac{4 M_N}{ m_\pi} \frac{f_{pp}^2 e \varepsilon}{k^2-m_\pi^2}  \frac{1}{m_{A^\prime}^2 - 2 \qa \cdot p_2} 
	\bar{u}(p_4) \gamma_5 u(p_1) \bar{u}(p_3) \gamma_5 (\slashed{p}_2 - \qa +M_N)\slashed \epsilon  u(p_2), \\
	M_5 &=& \frac{4 M_N}{ m_\pi} \frac{f_{pp}^2 e \varepsilon}{k^2-m_\pi^2}  \frac{1}{m_{A^\prime}^2 + 2 \qa \cdot p_3} 
	\bar{u}(p_3) \gamma_5 u(p_1) \bar{u}(p_4) \slashed{\epsilon} (\slashed{p}_3 + \qa +M_N)\gamma_5 u(p_2),\\
	M_6 &=& -\frac{4 M_N}{ m_\pi} \frac{f_{pp}^2 e \varepsilon}{l^2-m_\pi^2}  \frac{1}{m_{A^\prime}^2 + 2 \qa \cdot p_3} 
	\bar{u}(p_4) \gamma_5 u(p_1) \bar{u}(p_3) \slashed{\epsilon} (\slashed{p}_3 + \qa +M_N)\gamma_5 u(p_2),\\
	M_7 &=& \frac{4 M_N}{ m_\pi} \frac{f_{pp}^2 e \varepsilon}{k^2-m_\pi^2}  \frac{1}{m_{A^\prime}^2 + 2 \qa \cdot p_4} 
	\bar{u}(p_4) \gamma_5 u(p_2) \bar{u}(p_3) \slashed{\epsilon} (\slashed{p}_4 + \qa +M_N)\gamma_5 u(p_1),\\
	M_8 &=& -\frac{4 M_N}{ m_\pi} \frac{f_{pp}^2 e \varepsilon}{l^2-m_\pi^2}  \frac{1}{m_{A^\prime}^2 + 2 \qa \cdot p_4} 
	\bar{u}(p_3) \gamma_5 u(p_2) \bar{u}(p_4) \slashed{\epsilon} (\slashed{p}_4 + \qa +M_N)\gamma_5 u(p_1),\\
\end{eqnarray}
\end{widetext}
%
where the dark photon polarization is given by $\epsilon$. The squared amplitude is given by 
%
\begin{equation}
\vert \mathcal{M}_{\mathrm{pp}} \vert^2 = \sum_{s_1,s_2}\, \left\vert \sum_{i=1}^{8}\, M_i \right\vert^2, 
\end{equation}
%
where the first summation is over the incoming proton polarizations. Unfortunately, with the kinematical relations above, 
the squared amplitude does not yield a tidy expression for the spin-averaged squared matrix element. Our result 
contains over 200 terms, so we do not reproduce it here for reasons of convenience. However, we note that 
our result for $\vert \mathcal{M}_{\mathrm{pp}} \vert^2$ is symmetric under exchange of $k$ and $l$ as required.

 it does not seem to be possible to simplify the amplitude into an analytically usable form without making unjustified approximations, and so the final expression for the spin-averaged squared matrix element, $ \mathcal{M}^2_{pp}$, contains some two hundred and fifty terms, and thus we do not reproduce that result here. The pn calculation likewise results in a very large and unwieldy expression for $ \mathcal{M}^2_{pn}$. \footnote{At this point we diverge from Ref.~\cite{dent_etal12} for two reasons.  One, they use the kinematics as if the $A'$ were massless, as in the axion calculation, and independent of that choice, their squared matrix element does not respect the correct exchange symmetry, as can be seen by inspection of their Eq. (A.3), with their claim that only the coefficient $C_k$ appearing in that equation is nonzero.}

@@

\subsection{Streaming limit}
The first and simplest bound that may be obtained arises from assuming that all produced particles leave the supernova, carrying their energy with them. The constraint is derived simply by requiring that the energy loss through this cooling channel be roughly less than the cooling from neutrino emission; any greater, and it would have an observable effect on supernova cooling.

The actual quantity of interest therefore is the rate of energy emission through dark bosons. From the spin-summed squared amplitudes already found, this is obtained by integrating over the phase space, and adding a factor of energy of the emitted particle.
\beq
Q_i = \int (2\pi)^4 E_A \sum_{spins} \mathcal{M}^2_i f(p_1) f(p_2)\delta(p_1+p_2-p_3-p_4-\qa) d\Pi,
\eeq
where $d\Pi$ is the Lorentz-invariant phase space, $ E_A $ is the energy of the emitted boson, and $ f(p) $ are the initial occupation numbers. The nucleons in the core are comfortably non-degenerate and non-relativistic, so  the Pauli blocking factor is omitted and the Maxwell-Boltzmann distribution is used  
\beq
f(p) =  \frac{n_b}{2} (\frac{2 \pi}{M_N T})^{3/2} e^{-\frac{\textbf{p}^2} {2 M_N T}}.
\eeq

The integration is performed numerically to obtain $ Q_i$, which is the rate of energy emission per unit volume associated with either the $pn$ or $pp$ process. To obtain the dark gauge boson luminosity, it is assumed that production takes place in a core volume $V$ of radius 1 km, giving $ \mathcal{L}_A = V(Q_{pp} + Q_{pn}) $. The luminosity $\mathcal{L}_A $ is a complicated function of $ m_A $ and $ T $, but is simply proportional to the square of the parameter $ \varepsilon $.  Pulling that out of the expression gives a constraint 
\beq
\varepsilon^2 I_A(m_A, T) \le \mathcal{L}_\nu
\eeq
or a bound on the coupling of 
\beq 
\varepsilon \le \sqrt{\frac{4.1 \times 10^{37} MeV^2}{I_A(m_A, T)}} 
\eeq 
and the exclusion region is generated by varying the mass.

\subsection{Decay limit}
The constraint from the streaming limit above provides an upper bound on the allowed coupling (or lower bound on the excluded coupling) by considering the production on the $A'$. Naively it might be expected that this suffices, in that all higher couplings are excluded. However, in order to function as a cooling channel, enough of the produced dark gauge bosons must escape the supernova. As the coupling increases, so do processes that prevent the escape, and so the excluded region has an upper bound, above which the luminosity again drops below the neutrino luminosity. The first such limit may be found by considering decay of the dark bosons into Standard Model particles, and assuming that these SM particles are trapped in the supernova core and contribute nothing to the cooling. The dark boson has a typical lifetime of 
\beq
l = \frac{3 E_{A}}{N_{eff} m_A^2 \varepsilon^2},
\eeq
and so the fraction escaping the supernova before decaying is given by 
\beq
e^{r_{decay}/l} = e^{r_{decay} N_{eff} m_A^2 \varepsilon^2/(3 E_A)}.
\eeq
To take this into account, the above exponential factor is simply appended to the phase space integrand, and the calculation then proceeds as before. The limit is derived from the same equation, with the only complication being the fact that $ I_A $ is now a function of $ \varepsilon $, in addition to $ m_A, T $. This makes the numerical calculation slightly more complicated, but otherwise changes nothing of importance, with now the constraint a transcendental equation,  
\beq
\varepsilon \le \sqrt{\frac{4.1 \times 10^{37} MeV^2}{I_A(m_A, T, \varepsilon)}}.
\eeq

\subsection{Trapping limit}
The second constraint that produces an upper bound on the excluded region comes from considering trapping of dark bosons within the supernova. With a large enough coupling, the new particles will thermalize and then will be emitted from a spherical shell.  In this case the luminosity is given simply by the Steffan-Boltzmann law
\beq
\mathcal{L}_t  = 4\pi r^2 T_A^4 \sigma,
\eeq
where $ r $ is now the radius of the emitting shell and $ T_A $ its temperature. We can estimate $ r = 10$ km, since the density of the supernova drops drastically around that point. After taking that value for r, the bound on the luminosity translates into a bound on $ T_A $ 
\beq
T_A \le 9.586\rm\ MeV.
\eeq
That bound can then be translated into the desired bound on the coupling as a function of mass by assuming that the particles are emitted from an optical depth $ \tau = 2/3 $, and finding the temperature that corresponds to that optical depth. This is a somewhat involved calculation.
First, one needs a model for the density and temperature in the supernova. Following \cite{dent_etal12}, we assume 
\bea
\rho &=& \rho_p \left(\frac{R}{r}\right)^n,\\
T &=& T_R \left(\frac{\rho(r)}{\rho_R}\right)^{1/3},
\eea
with $  \rho_p  = 3\times10^{14}\ {\rm g/cm}^3, T_R = 30$ MeV, and taking $ n = 5 $. The optical depth is given by 
\beq
\tau = \int_{r_x}^{\infty} \kappa \rho dr,
\eeq
where $ \kappa$ is the opacity.  To find the opacity, we start from the reduced mean Rosseland opacity 
\beq
\frac{1}{\kappa \rho} = \int_{m_x}^{\infty} \frac{15}{4 \pi^4 T^5} \frac{E_A^2 e^{E_A/T} \sqrt{E_A^2 - m_A^2}}{(e^{E_A/T}-1)^2} l_A dE_A,
\eeq
where $ l_A $ is the mean free path.

@@

\subsection{Phase Space Integration}
In every case calculating the bound requires integrating an expression involving the long and complicated Bremsstrahlung amplitude, which unfortunately could not be done analytically without making unjustified approximations. The integration was consequently performed numerically using the Monte Carlo routines provided by the Cuba library. Employing kinematic relations did not seem to reduce the complexity of the problem, so in the interests of reducing the number of possible mistakes the kinematics were done by brute force. We integrated over the entire phase space explicitly in terms of the eleven variables of $ \vec{p_1}, \vec{p_2}, \vec{p_3}, \hat{\qa} $, then fixed $ p_4 $ using the 3-momentum part of the delta-function and used an unpleasantly complicated equation obtained from the energy delta-function to fix the remaining one free momentum amplitude - in our case we chose this to be $ \qa $, the momentum of the emitted boson. At each step points resulting in unphysical configurations were discarded, and the remaining points were then plugged in to the appropriate amplitude, and then that result used as the integrand. The numerical integration routine used was the \textit{Suave} method provided by CUBA, which combines importance sampling and adaptive subdivision; this method seemed to provide the best compromise between computation time and error.

@@

We then scanned over the desired range of values for $ M_A $ and output the three limits - trapping, decay, and streaming - at each point. The complexity of the integration was such that errors remained relatively large even with highest practicable number of samples.

The calculation is dependent on a number of constants. We chose the following values as roughly typical of supernovae: T, the temperature of the supernova core, was set to 30 MeV; r, the radius of the supernova core, to 1 km; $ R $, the distance at which decay products are trapped, to 10 km; $ \mathcal{L} $, the luminosity bound, to $ 10^{53} erg/s $; and $ \n_b $, the baryon number density in the supernova interior, to $ 1.79*10^{38} cm^{-3} $.

@@

%------------ APPENDIX
\appendix
\label{appendix:details}

\section{Bremsstrahlung diagrams}
	Each of the $p-p-\pi $ vertices contribute a factor of $g_{pp}$,  the pseudoscalar pion-nucleon coupling. $f_{pp}$, the pseudovector pion-nucleon coupling, is related by $g_{pp} = \frac{2M_N}{m_\pi} f_{pp} $. The $ p-p-A' $ vertex contributes $e\varepsilon \gamma_\mu$, where $\varepsilon$ is the mixing strength between the photon and dark photon. Labeling the contributions according to the Feynman diagrams in Fig. , the matrix elements are where the exchange momenta are defined by $ k = p_2-p_4 $ and $ l = p_2-p_3 $, and the dark photon's polarization is given by $\epsilon$. 
	For $pn$ scattering, four of the diagrams are simply the same as those given above. The only new Feynman diagram comes from bremsstrahlung off the internal pion, and results in  
	 \beq
	 M'_5 = \frac{4 M_N}{ m_\pi} \frac{f_{pn}^2 e \varepsilon}{l^2-m_\pi^2
	 }  \frac{1}{(l-\qa)^2 - m_\pi^2} \bar{u}(p_4) \gamma_5 u(p_1) \bar{u}(p_3) u(p_2) (\qa - 2l)\cdot \varepsilon
	 \eeq
	 Note that these matrix elements agree with Ref.~\cite{dent_etal12}.
\acknowledgments

>>>>

\arz{Cameron, can you double check this. I think there were some errors in the momenta and in the relative 
signs of the diagrams. I tried to fixe them.}
The eight diagrams contribute the following eight terms to the pp amplitude, 
%
\begin{widetext}
\begin{eqnarray}
	M_1 &=& \frac{4 M_N}{ m_\pi} \frac{f_{pp}^2 e \varepsilon}{k^2-m_\pi^2}  \frac{1}{m_{A^\prime}^2 - 2 \qa \cdot p_1} 
	\bar{u}(p_4) \gamma_5 u(p_2) \bar{u}(p_3) \gamma_5 (\slashed{p}_1 - \qa +M_N)\slashed \epsilon  u(p_1), \\
	M_2 &=& -\frac{4 M_N}{ m_\pi} \frac{f_{pp}^2 e \varepsilon}{l^2-m_\pi^2}  \frac{1}{m_{A^\prime}^2 - 2 \qa \cdot p_1} 
	\bar{u}(p_3) \gamma_5 u(p_2) \bar{u}(p_4) \gamma_5 (\slashed{p}_1 - \qa +M_N)\slashed{\epsilon} u(p_1),\\
	M_3 &=& \frac{4 M_N}{ m_\pi} \frac{f_{pp}^2 e \varepsilon}{l^2-m_\pi^2}  \frac{1}{m_{A^\prime}^2 - 2 \qa \cdot p_2} 
	\bar{u}(p_3) \gamma_5 u(p_1) \bar{u}(p_4) \gamma_5 (\slashed{p}_2 - \qa +M_N)\slashed \epsilon  u(p_2), \\
	M_4 &=& -\frac{4 M_N}{ m_\pi} \frac{f_{pp}^2 e \varepsilon}{k^2-m_\pi^2}  \frac{1}{m_{A^\prime}^2 - 2 \qa \cdot p_2} 
	\bar{u}(p_4) \gamma_5 u(p_1) \bar{u}(p_3) \gamma_5 (\slashed{p}_2 - \qa +M_N)\slashed \epsilon  u(p_2), \\
	M_5 &=& \frac{4 M_N}{ m_\pi} \frac{f_{pp}^2 e \varepsilon}{k^2-m_\pi^2}  \frac{1}{m_{A^\prime}^2 + 2 \qa \cdot p_3} 
	\bar{u}(p_3) \gamma_5 u(p_1) \bar{u}(p_4) \slashed{\epsilon} (\slashed{p}_3 + \qa +M_N)\gamma_5 u(p_2),\\
	M_6 &=& -\frac{4 M_N}{ m_\pi} \frac{f_{pp}^2 e \varepsilon}{l^2-m_\pi^2}  \frac{1}{m_{A^\prime}^2 + 2 \qa \cdot p_3} 
	\bar{u}(p_4) \gamma_5 u(p_1) \bar{u}(p_3) \slashed{\epsilon} (\slashed{p}_3 + \qa +M_N)\gamma_5 u(p_2),\\
	M_7 &=& \frac{4 M_N}{ m_\pi} \frac{f_{pp}^2 e \varepsilon}{k^2-m_\pi^2}  \frac{1}{m_{A^\prime}^2 + 2 \qa \cdot p_4} 
	\bar{u}(p_4) \gamma_5 u(p_2) \bar{u}(p_3) \slashed{\epsilon} (\slashed{p}_4 + \qa +M_N)\gamma_5 u(p_1),\\
	M_8 &=& -\frac{4 M_N}{ m_\pi} \frac{f_{pp}^2 e \varepsilon}{l^2-m_\pi^2}  \frac{1}{m_{A^\prime}^2 + 2 \qa \cdot p_4} 
	\bar{u}(p_3) \gamma_5 u(p_2) \bar{u}(p_4) \slashed{\epsilon} (\slashed{p}_4 + \qa +M_N)\gamma_5 u(p_1),
\end{eqnarray}
\end{widetext}
%
where the dark photon polarization is given by $\epsilon$. The squared amplitude is given by 
%
\begin{equation}
\vert \mathcal{M}_{\mathrm{pp}} \vert^2 = \sum_{s_1,s_2}\, \left\vert \sum_{i=1}^{8}\, M_i \right\vert^2, 
\end{equation}
%
where the first summation is over the incoming proton polarizations. These expressions are identical to those 
given in Ref.~\cite{dent_etal12}; however, they do not simplify significantly if the correct kinematics are used. 
Unfortunately, with the kinematical relations above, the squared amplitude does not yield a tidy expression 
for the spin-averaged squared matrix element. Our result 
contains over 200 terms, so we do not reproduce it here for reasons of convenience. However, we note that 
our result for $\vert \mathcal{M}_{\mathrm{pp}} \vert^2$ is symmetric under exchange of $k$ and $l$ as required.

The pn process contain four diagrams analogous to the the diagrams for the pp process (there are only four, of course, 
because the neutrons cannot radiate the $A'$). The new diagram that is relevant in the pn process is shown in 
Fig.~\ref{fig:npdiagram} and yields a contribution of 
%
\begin{widetext}
\begin{equation}
M'_5 = \frac{4 M_N}{ m_\pi} \frac{f_{pn}^2 e \varepsilon}{l^2-m_\pi^2}  \frac{1}{(l-\qa)^2 - m_\pi^2} 
\bar{u}(p_3) \gamma_5 u(p_1) \bar{u}(p_4) \gamma_5 u(p_2) (\qa - 2l)\cdot \varepsilon.
\end{equation}
\end{widetext}
%
The pn processes likewise yield a squared amplitude, 
$\vert \mathcal{M}_{\mathrm{pn}} \vert^2$ that is unwieldy, so we do not give the 
the full expression here.

@@

%------------------------------------- STREAMING CONSTRAINT
\subsection{The Streaming Limit of the Energy Loss Rate}

The first and simplest bound that may be obtained arises from assuming that all $A'$ particles produced in the supernova core 
leave the supernova, carrying their energies with them. The constraint can be derived simply by 
requiring that the energy loss through this cooling channel be less than the cooling from neutrino emission; 
any greater, and it would have an observable effect on supernova cooling. This calculation yields values of 
$\varepsilon$ above which cooling through $A'$ production is too rapid to be consistent with SN1987A. 
We will consider modifications to this bound from $A'$ trapping and decay in subsequent subsections.

The quantity of interest is the rate of energy emission through dark gauge bosons. 
From the spin-summed, squared amplitudes described in the previous subsection, 
the energy emission rate is obtained by integrating over the phase space, 
and adding a factor of the energy of the emitted particle. To be specific, the 
energy emission rate per unit volume is 
%
\begin{widetext}
\beq
\label{eq:rate1}
Q_i = (2\pi)^4 \int\, E_{\mathrm{A}'} \, \sum_{s_1,s_2}\, \vert \mathcal{M}_i \vert^2 f(p_1) f(p_2)\delta^{(4)}(p_1+p_2-p_3-p_4-\qa)\, d\Pi,
\eeq
\end{widetext}
%
where $d\Pi$ is the Lorentz-invariant phase space, 
$ E_{\mathrm{A}'}$ is the energy of the emitted $A'$ boson, 
$f(p)$ are the phase-space densities of the incoming nucleons, and 
the index $i$ refers to either the pp or pn processes. 
The nucleons in the core are comfortably non-degenerate and non-relativistic, 
so the Pauli blocking factor is omitted from Eq.~(\ref{eq:rate1}) and we take 
all nucleons to have a Maxwell-Boltzmann momentum distribution, so 
\beq
f(p) =  \frac{n_b}{2} \left( \frac{2 \pi}{M_N T} \right)^{3/2} \exp \left(-\frac{p^2}{2 M_N T} \right).
\eeq
%
We chose a baryon number density of $n_{\mathrm{b}} \approx 1.8 \times 10^{38}\, \mathrm{cm}^{-3}$ and a 
core supernova temperature of $T = 30\, \mathrm{MeV}$, both of which are typical choices and thought to 
be representative of supernova cores. \arz{Need to put a citation here for the properties of supernova cores. 
I can probably find something in the literature, but please put something here if you have a citation.}

We performed the phase space integrals using the Monte Carlo routines in the {\tt CUBA} library. 
\arz{There needs to be some reference, citation, or some other information give about this library here.} 
We integrated over the momenta $\vec{p}_1$, $\vec{p}_2$, $\vec{p}_3$, and the direction of the 
three-momentum of the radiated boson $\hat{q}_{\mathrm{A}'}$, after fixing $\vec{p}_4$ 
and the magnitude of $\vec{q}_{\mathrm{A}'}$ using the delta functions. 
\arz{Cameron, can you double check this to be sure that it is correct? I rephrased it a bit, 
but I found the earlier phrasing to be a bit ambiguous and I can't remember which choice 
you made.} We used the {\tt suave} method provided within {\tt CUBA}, which combines 
importance sampling and adaptive subdivision, as this method provided the best compromise 
between accuracy and run-time for this particular application.

To obtain the dark gauge boson luminosity from $Q_{\mathrm{pp}}$ and $Q_{\mathrm{pn}}$, 
we assumed that $A'$ production takes place in a stellar core of radius $\sim 1$~km, so that 
the total luminosity of $A'$ is 
%
\begin{equation}
L_{\mathrm{A}'} = V (Q_{\mathrm{pp}} + Q_{\mathrm{pn}})
\end{equation}
%
where $V$ is the volume of the sphere. Clearly, $L_{\mathrm{A}'}$ is proportional to $\varepsilon^2$, 
so that writing $L_{\mathrm{A}'} = \varepsilon^2 I_{\mathrm{A}'}(m_{\mathrm{A}'},T)$ and 
$L_{\nu}$ as the neutrino luminosity, we can write the constraint as 
%
\begin{equation}
\varepsilon \lesssim \sqrt{\frac{L_{\nu}}{I_{\mathrm{A}'}(m_{\mathrm{A}'}, T)}}.
\end{equation}
%
Following SOMEBODAY \arz{We need a citation here.}, we show a constraint taking 
$L_{\nu} \approx 10^{53}\, \mathrm{erg/s} \approx 4 \times 10^{37}\, \mathrm{MeV}^2$.

%---------------------------------- DECAY LIMIT
\subsection{Decay limit}
The constraint from the streaming limit above provides an upper bound on the allowed coupling (or lower bound on the excluded coupling) by considering the production on the $A'$. Naively it might be expected that this suffices, in that all higher couplings are excluded. However, in order to function as a cooling channel, enough of the produced dark gauge bosons must escape the supernova. As the coupling increases, so do processes that prevent the escape, and so the excluded region has an upper bound, above which the luminosity again drops below the neutrino luminosity. The first such limit may be found by considering decay of the dark bosons into Standard Model particles, and assuming that these SM particles are trapped in the supernova core and contribute nothing to the cooling. The dark boson has a typical lifetime of 
\beq
l = \frac{3 E_{A}}{N_{eff} m_A^2 \varepsilon^2},
\eeq
and so the fraction escaping the supernova before decaying is given by 
\beq
e^{r_{decay}/l} = e^{r_{decay} N_{eff} m_A^2 \varepsilon^2/(3 E_A)}.
\eeq
To take this into account, the above exponential factor is simply appended to the phase space integrand, and the calculation then proceeds as before. The limit is derived from the same equation, with the only complication being the fact that $ I_A $ is now a function of $ \varepsilon $, in addition to $ m_A, T $. This makes the numerical calculation slightly more complicated, but otherwise changes nothing of importance, with now the constraint a transcendental equation,  
\beq
\varepsilon \le \sqrt{\frac{4.1 \times 10^{37} MeV^2}{I_A(m_A, T, \varepsilon)}}.
\eeq

%---------------------------------- TRAPPING	
\subsection{Trapping limit}
The second constraint that produces an upper bound on the excluded region comes from considering trapping of dark bosons within the supernova. With a large enough coupling, the new particles will thermalize and then will be emitted from a spherical shell.  In this case the luminosity is given simply by the Steffan-Boltzmann law
\beq
\mathcal{L}_t  = 4\pi r^2 T_A^4 \sigma,
\eeq
where $ r $ is now the radius of the emitting shell and $ T_A $ its temperature. We can estimate $ r = 10$ km, since the density of the supernova drops drastically around that point. After taking that value for r, the bound on the luminosity translates into a bound on $ T_A $ 
\beq
T_A \le 9.586\rm\ MeV.
\eeq
That bound can then be translated into the desired bound on the coupling as a function of mass by assuming that the particles are emitted from an optical depth $ \tau = 2/3 $, and finding the temperature that corresponds to that optical depth. This is a somewhat involved calculation.
First, one needs a model for the density and temperature in the supernova. Following \cite{dent_etal12}, we assume 
\bea
\rho &=& \rho_p \left(\frac{R}{r}\right)^n,\\
T &=& T_R \left(\frac{\rho(r)}{\rho_R}\right)^{1/3},
\eea
with $  \rho_p  = 3\times10^{14}\ {\rm g/cm}^3, T_R = 30$ MeV, and taking $ n = 5 $. The optical depth is given by 
\beq
\tau = \int_{r_x}^{\infty} \kappa \rho dr,
\eeq
where $ \kappa$ is the opacity.  To find the opacity, we start from the reduced mean Rosseland opacity 
\beq
\frac{1}{\kappa \rho} = \int_{m_x}^{\infty} \frac{15}{4 \pi^4 T^5} \frac{E_A^2 e^{E_A/T} \sqrt{E_A^2 - m_A^2}}{(e^{E_A/T}-1)^2} l_A dE_A,
\eeq
where $ l_A $ is the mean free path.

@@

\subsection{Phase Space Integration}

In every case

@@

@@

\acknowledgments
